Supplementary materials

### 2018-07-23 WholePlotTrajectories.Rmd
annex

This variability of protocols in time raised methodological issues as vernacular names usually correspond to different botanical species. 
It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_3\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:

\begin{align}
M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_3\big]\Big) \nonumber
\end{align}

@@

While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I). For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)). The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.87$ and $\rho_{spearman}^{functional}=0.90$ respectively). The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.

@@

The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna.

@@

Our study revealed community recovery through the combination of deterministic processes driving their convergence in functional composition, and dispersal limitation maintaining their divergence in taxonomic composition.
The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evennes that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functoinal composition.
The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pionneers that prevented in the first times the establishment of more diverse long-lived pionneers, recruited otherwise below the intensity threshold.
The resilience of tropical forests proved consistent although several decades-long. 
Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].
As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.

>>>>

This variability of protocols in time raised methodological issues as vernacular names usually correspond to different botanical species. 
It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_V\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:

\begin{align}
M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_V\big]\Big) \nonumber
\end{align}

@@

While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I). For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)). The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.87$ and $\rho_{spearman}^{functional}=0.90$ respectively). The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.

@@

The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
Besides, the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna.

@@

Our study revealed community recovery through the combination of deterministic processes driving their convergence in functional composition, and dispersal limitation maintaining their divergence in taxonomic composition.
The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evennes that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functoinal composition.
The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pionneers that prevented in the first times the establishment of more diverse long-lived pionneers, recruited otherwise below the intensity threshold.
The resilience of tropical forests proved consistent although several decades-long. 
Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].

### 2018-07-23 WholePlotTrajectories_SuppMat.Rmd
annex

---
title: "Supplementary materials"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric MARCON
    authsuperscript: 1
  - name: Bruno HERAULT
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB (Institut National Ploytechnique Félix Houphoüet Boigny) 
    address: Yamoussoukro, Ivory Coast
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: http://www.ecofog.gf/spip.php?article47
journalinfo: Hal 00679993
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings}
lang: english
output: 
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::gitbook: default
  bookdown::word_document2: default
---

@@

>>>>

---
title: "Supplementary materials"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric MARCON
    authsuperscript: 1
  - name: Bruno HERAULT
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB (Institut National Ploytechnique Félix Houphoüet Boigny) 
    address: Yamoussoukro, Ivory Coast
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: http://www.ecofog.gf/spip.php?article47
journalinfo: Hal 00679993
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings}
lang: english
output: 
  bookdown::gitbook: default
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
---

@@

The biodiversity estimator is based on the association probability between vernacular names and botanical species. 
The association probability is estimated by a Dirichlet trial based on the vernacular/botanical association frequency observed in the whole inventory.
This association probability then allows generating theoretical inventories with complete botanical identification with the assignment of botanical species to trees with only a vernacular name.
The repetition of the replacement process produces a set of theoretical inventories which diversities gives the probability distribution of the initial inventory diversity.
